### 2021-01-31 papers/jin-thesis/Chapters/1Introduction.tex
Edited thesis
Edited thesis by adding refs and adjusted transition matrix part

	Neurodegenerative diseases, such as Alzheimer's and Huntinton's disease, have consistently been one of the leading causes of death in the United States. Scientists all over the world struggle to find a solution for these diseases and continue to hope for a breakthrough in science in attempt to prevent or even cure these fatal diseases. Although this breakthrough has yet to come, research suggests that motor protein failure is a main cause of cell degeneration. More specifically, defects in the motor protein dynein has been linked to neurodegeneration due to the cell's funtional dependence on dynein and its mechanical properties. However, unlike dynein's motor protein siblings, dynein possesses a unique structure causing it to stochasitcally move around the cell and be unpredicatable in nature. Thus, fully understanding the mechanisms of dynein can allow us to further conclude causes of its failure, putting us one step closer to discovering the major breakthrough for neurodegenerative research. In this paper, we propose a model of dynein that uses Monte Carlo methods and a Brownian dynamics simulation in order to reproduce the steping behavior observed from experiment.

@@

Cells are complicated. Despite their fatal degeneracies, cells have the ability to move, perfectly divide, and spatially organize themselves. They can do so because tiny, two-legged molecular proteins, named \textit{motor proteins}, carry cargo containing information that outlines the blueprint for different functions within the cell. These motor proteins consists of kinesin, myosin, and dynein. They can travel all around the cell on a cytoskeleton composed of microtubules (MT) that act as a walkway for the proteins. A depiction of dynein walking along the microtubule is shown above in Figure (\ref{fig:ArtDynein}). Microtubules are constructed from negatively charged $\alpha$- and positively charged $\beta$-tubulin dimers that polarizes the microtubule depending on which tubulin is left hanging at the ends. In a eukaryotic cell, microtubules typically stem outwards from the cell's nucleus to the cell walls where the $\beta$ tubulin is exposed, generating a positive end that pulls kinesin and myosin. However, dynein opposes kinesin and myosin by walking towards the nucleus cell body where the MT is negatively charged. This attraction to both ends drives the directed stepping motion for each motor protein and is shown in Figure (\ref{fig:transport}).

@@

Both kinesin and myosin exhibit hand-over-hand stepping, where their ``legs" take turns alternating steps like a human. Due to dynein's massive size compared to its motor protein siblings, dynein does not step uniformly but rather inchworms its ``feet" like a drunk human. This unique property makes dynein difficult to study when analyzing possible sources of molecular failure, but utters the importance of understanding the physical structure of the entire motor protein.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\columnwidth]{Figures/dynein_on_MT.png}
	\caption[Structure of Dynein]{\textbf{Structure of Dynein.} Illustration of cytoplasmic dynein's heavy chain subunits bounded to the microtubule. The large size of the AAA+ rings are blocking the view of the independent linkers connecting to the tail. This rendition similarly illustrates dynein’s structure as a domain-rod system. \cite{TheTrappistArt}}
	\label{fig:structure}
\end{figure}

Dynein's crystal structure has been intensively studied because of the pivotal role it plays in dyneins unpredictable stepping nature. Its inability to step uniformly is largely due to the large lengths between each major domain site, as the linker and stalk are about a factor of a half larger than the diameter of the AAA+ ring. This is unique when commpared to kinesin and myosin, as their motor domains are much closer to the MT causing their center of mass to keep the protein controlled and steady when stepping. Despite dynein's large structure, dynein can still achieve processive stepping towards the negative end of the MT due to the interaction of adenosine triphosphate (ATP) with the AAA+ domains, thus is why the AAA+ domains are referred to as the motor domains.

@@

Dynein's ability to step is entirely ATP driven under the process of ATP hydrolysis with its motor domains. The ATP provides dynein chemical energy which converts into mechanical energy for the dynein to transition between differenet conformational states during a step. Currently, there are many models being explored that discusses how dynein uses the ATP to step, such as the powerstroke model or the winch model, but this paper will be influenced by the more accepted powerstroke model. For either case, it is best to define an axis on the MT for where dynein can step.

@@

Simulations that model dynein's step exists, but all of them are used to investigate the ATP hydrolysis mechanism during the mechanochemical cycle. Because of this, these simulations suffer from numerous assumptions made about the physical aspects of dynein's step. (\textit{Should reference a simulation here and be more detailed}) Some assume that dynein can only step in increments of 8nm or that the rebinding process is constrained to a certain time frame. These limitations make it impossible for the models to replicate unique discoveries about dyneins stepping, such as Yildiz's correlation from above. A perfect simulation ought to lift these physical constraints and accurately reproduce experimental data.

>>>>

	Neurodegenerative diseases, such as Alzheimer's and Huntinton's disease, have consistently been one of the leading causes of death in the United States. Scientists all over the world struggle to find a solution for these diseases and continue to hope for a breakthrough in science in attempt to prevent or even cure these fatal diseases. Although this breakthrough has yet to come, research suggests that motor protein failure is a main cause of cell degeneration. More specifically, defects in the motor protein dynein has been linked to neurodegeneration due to the cell's funtional dependence on dynein and its mechanical properties. However, unlike dynein's motor protein siblings, dynein possesses a unique structure causing it to stochasitcally move around the cell and be unpredicatable in nature. Thus, fully understanding the mechanisms of dynein can allow us to further conclude causes of its failure, putting us one step closer to discovering the major breakthrough for neurodegenerative research. In this paper, we propose a model of dynein that uses Monte Carlo methods and a Brownian dynamics simulation to reproduce the steping behavior observed from experiment. \textit{ADD CLARIFIER}

@@

Cells are complicated. They have the ability to move, perfectly divide, and spatially organize themselves when they are functional, but can also be responsible for death when they are not. They can function because tiny, two-legged molecular proteins, named \textit{motor proteins}, carry cargo with information around the cell. These motor proteins consists of kinesin, myosin, and dynein. They can travel all around the cell on a cytoskeleton composed of microtubules (MT) that act as a walkway for the proteins. A depiction of dynein walking along the microtubule is shown above in Figure (\ref{fig:ArtDynein}). Microtubules are constructed from negatively charged $\alpha$- and positively charged $\beta$-tubulin dimers that polarizes the microtubule depending on which tubulin is left hanging at the ends. In a eukaryotic cell, microtubules typically stem outwards from the cell's nucleus to the cell walls where the $\beta$ tubulin is exposed, generating a positive end that pulls kinesin and myosin. However, dynein opposes kinesin and myosin by walking towards the nucleus cell body where the MT is negatively charged. This attraction to both ends drives the directed stepping motion for each motor protein and is shown in Figure (\ref{fig:transport}).

@@

Both kinesin and myosin exhibit hand-over-hand stepping, where their ``legs" take turns alternating steps like a human. Due to dynein's massive size compared to its motor protein siblings, dynein does not step uniformly but rather inchworms its ``feet" like a drunk human. This unique property makes dynein difficult to study when analyzing possible sources of molecular failure, but illustrates the importance of understanding the physical structure of the entire motor protein.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\columnwidth]{Figures/dynein_on_MT.png}
	\caption[Structure of Dynein]{\textbf{Structure of Dynein.} Illustration of cytoplasmic dynein's heavy chain subunits bounded to the microtubule. The large size of the AAA+ rings are blocking the view of the independent linkers connecting to the tail. This rendition depicts dynein’s structure as a domain-rod system. \cite{TheTrappistArt}}
	\label{fig:structure}
\end{figure}

Dynein's crystal structure has been rigorously studied because of the pivotal role it plays in dyneins unpredictable stepping nature. Its inability to step uniformly is largely due to the large lengths between each major domain site, as the linker and stalk are about a factor of a half larger than the diameter of the AAA+ ring. This is unique when commpared to kinesin and myosin, as their motor domains are much closer to the MT causing their center of mass to keep the protein controlled and steady when stepping. Despite dynein's large structure, dynein can still achieve processive stepping towards the negative end of the MT due to the interaction of adenosine triphosphate (ATP) with the AAA+ domains, thus is why the AAA+ domains are referred to as the motor domains.

@@

Dynein's ability to step is entirely ATP driven under the process of ATP hydrolysis with its motor domains. The ATP provides dynein chemical energy which converts into mechanical energy for the dynein to transition between differenet conformational states during a step. Currently, there are many models being explored that discuss how dynein uses the ATP to step, such as the powerstroke model or the winch model, but this paper will be influenced by the more accepted powerstroke model. For either case, it is best to define an axis on the MT for where dynein can step.

@@

Simulations that model dynein's step exists, but all of them are used to investigate the ATP hydrolysis mechanism during the mechanochemical cycle. These simulations suffer from numerous assumptions made about the physical aspects of dynein's step. (\textit{Should reference a simulation here and be more detailed}) Some assume that dynein can only step in increments of 8nm or that the rebinding process is constrained to a certain time frame. These limitations make it impossible for the models to replicate unique discoveries about dyneins stepping, such as Yildiz's correlation from above. A perfect simulation ought to lift these physical constraints and accurately reproduce experimental data.

### 2021-01-31 papers/jin-thesis/Chapters/3Methods.tex
Edited thesis by adding refs and adjusted transition matrix part

The entire model and simulation code can be found here on \url{https://github.com/kiatvonj/dynein_walk}, where the model and simulation were coded using both Python and C++. The both bound Monte Carlo was coded on Python due to the calculations being less computationally rigorous and time independent, while the one bound Brownian dynamics was coded on C++ to utilize the faster computation speed for the time  dependent Brownian motion.

\section{Simulation}
%\textit{Picking Random angles to create random distribution of dynein configurations. Make sure to clarify that the simulation is split into two parts, the monte carlo both bound and brownian one bound. It is done this way in order for MC to generate ensemble of data rather quickly.}\\
%\textit{Include a flow chart.}
Since we want to generate an ensemble of steps, we use the Monte Carlo method to simulate a single step at a time, as opposed to a walk. This allows us to simplify the code into two respective parts (the Monte Carlo both bound and the Brownian dynamics one bound) that cycles repeatedly after the model takes a step. The simulation starts in the both bound state, where the Monte Carlo randomly picks a both bound configuration and calculates the total energy. If the configuration successfully unbinds, then we use the positions of the domains as initial conditions for the one bound simulation and execute the Brownian dynamics. Once the one bound dynein diffuses onto the microtubule and rebinds, the simulation repeats the cycle and generates a completely different both bound configuration for another step. However, this method comes with a caveat: we must initialize the distance between the binding domains, $L$, in order to ensure that our ensemble covers the whole sample space of possible both bound configurations. This will force the measured probabilities to be a function of $L$ and will be addressed later in the Data Analysis section (Section \ref{sec:DataAna}). The algorithm for the both bound process is shown below:

@@

\begin{figure}[hbt!]
	\centering
	\includegraphics[width=0.3\columnwidth]{../../plots/burgess-model-figure.pdf}
	\includegraphics[width=0.5\columnwidth]{../../plots/grotjahn-model-figure.pdf}%
	\caption[Experimental Pictures of Dynein for Model Parameters]{\textbf{Experimental Pictures of Dynein for Model Parameters} \textit{Left:} Micrograph of both bound dynein with model domains to identify post-stroke equilbrium angles \cite{burgess2003dynein}.  \textit{Right:} Cryo-electron tomograph of one bound dynein for pre-stroke equilbrium angles \cite{grotjahn}.} 
	\label{fig:ParamsPics}
\end{figure}

@@

\section{Data Analysis} \label{sec:DataAna}
As mentioned earlier, the simulation can only collect statistics for a given initial binding domain distance, $L$. This limits the probability data of landing at a final position to be a function of $L$. However, the unsigned distance is insignificant when calculating the step length of either a trailing or leading step. This promotes the new naming scheme of using initial displacement, $x_i$, and final displacement $x_f$, when identifying the distance between the binding domains before and after the step. The displacement is the $x$ position of the unbounded binding domain during the step, where the origin is defined at the bounded binding domain. This implies a negative initial displacement for a trailing step and a positive initial displacement for a leading step.

@@

\[
	p(x_f|x_i)=\frac{p(x_i,x_f)}{p(x_i)},
\]
\begin{equation}
	p(x_i,x_f)=p(x_f|x_i)p(x_i).
\end{equation}
Now, we only need to find the probability density of initial displacements, $p(x_i)$. Since the initial displacement is simply a signed distance $L$, we can use the probability of unbinding for a trailing or leading step as a function of $L$ from Equation (\ref{eqn:ProbTrail}) to generate $p(x_i)$. In otherwords, Equation (\ref{eqn:ProbTrail}) gives us $p(x_i|L_i)$. We can relate the two by using Bayes' Theorem:

@@

\begin{equation}
	p(x_i)=p(x_i|L_i)p(L_i).
\end{equation}
Now, the final step is to solve for the probability density of initial distance, $p(L_i)$. We can do so by solving a Detailed Balance Equation, where we define a transition matrix, $T=p(L_f|L_i)$. This equation will have the form,
\begin{equation}
	p(L_i)=T^np^0(L_i),
\end{equation}
where $T$ is raised to the $n$ number of steps to signify equilibrium for the probability of all distances $L$. $n$ is chosen to be arbitrarily large to ensure equilibrium after enough steps (??? FIXME this reasoning needs to be clearer. Ask Roundy.) To solve for $T$, we can use the previous probability distributions $p(x_f|x_i)$ and $p(x_i|L_i)$ to convert the probabilities to distances, as shown below:
\begin{align}
	T&=I(L_f|x_i)p(x_f|x_i)p(x_i|L_i)\\
	&=p(L_f|x_i)p(x_i|L_i)\\
	&=p(L_f|L_i),
\end{align}
where $I(L_f|x_i)$ is the identity matrix for converting $x_f$ into $L_f$. We have now solved for all the unknowns to generate a joint probability distribution of initial and final displacements, $p(x_i,x_f)$.

>>>>

The entire model and simulation code can be found at \url{https://github.com/kiatvonj/dynein_walk}, where the model and simulation were coded using both Python and C++. The both bound Monte Carlo was coded on Python due to the calculations being less computationally rigorous and time independent, while the one bound Brownian dynamics was coded on C++ to utilize the faster computation speed for the time  dependent Brownian motion.

\section{Simulation}
%\textit{Picking Random angles to create random distribution of dynein configurations. Make sure to clarify that the simulation is split into two parts, the monte carlo both bound and brownian one bound. It is done this way in order for MC to generate ensemble of data rather quickly.}\\
%\textit{Include a flow chart.}
To generate an ensemble of steps, we use the Monte Carlo method to simulate a single step at a time, as opposed to a walk. This simplifies the code into two parts (the Monte Carlo both bound and the Brownian dynamics one bound) that cycle repeatedly after the model takes a step. The simulation starts in the both bound state, where the Monte Carlo randomly picks a both bound configuration and calculates the total energy. If the configuration successfully unbinds, then we use the positions of the domains as initial conditions for the one bound simulation and execute the Brownian dynamics. Once the one bound dynein diffuses onto the microtubule and rebinds, the simulation repeats the cycle and generates a completely different both bound configuration for another step. However, this method comes with a caveat: we must initialize the distance between the binding domains, $L$, in order to ensure that our ensemble covers the whole sample space of possible both bound configurations. This forces the measured probabilities to be a function of $L$ and will be addressed later in the Data Analysis section (Section \ref{sec:DataAna}). The algorithm for the both bound process is shown below:

@@

\begin{figure}[hbt!]
	\centering
	\includegraphics[width=0.3\columnwidth]{../../plots/burgess-model-figure.pdf}
	\includegraphics[width=0.5\columnwidth]{../../plots/grotjahn-model-figure.pdf}%
	\caption[Experimental Pictures of Dynein for Model Parameters]{\textbf{Experimental Pictures of Dynein for Model Parameters} \textit{Left:} Micrograph of both bound dynein with model domains to identify post-stroke equilbrium angles \cite{Burgess2003}.  \textit{Right:} Cryo-electron tomograph of one bound dynein for pre-stroke equilbrium angles \cite{grotjahn2018cryo}.} 
	\label{fig:ParamsPics}
\end{figure}

@@

\section{Data Analysis} \label{sec:DataAna}
As mentioned earlier, the simulation can only collect statistics for a given initial binding domain distance, $L$. This limits the probability data of landing at a final position to be a function of $L$. However, we can not calculate the step length for a trailing or leading step with an unsigned distance. This promotes the new naming scheme of using initial displacement, $x_i$, and final displacement $x_f$, when identifying the distances between the binding domains before and after the step. The displacement is the $x$ position of the unbounded binding domain during the step, where the origin is defined at the bounded binding domain. This implies a negative initial displacement for a trailing step and a positive initial displacement for a leading step.

@@

\[
	p(x_f|x_i)=\frac{p(x_i,x_f)}{p(x_i)},
\]
\begin{equation}
	p(x_i,x_f)=p(x_f|x_i)p(x_i).
\end{equation}
Now, we only need to find the probability density of initial displacements, $p(x_i)$. Since the initial displacement is simply a signed distance $L$, we can use the probability of unbinding for a trailing or leading step as a function of $L$ from Equation (\ref{eqn:ProbTrail}) to generate $p(x_i)$. In other words, Equation (\ref{eqn:ProbTrail}) gives us $p(x_i|L_i)$. We can relate the two by using Bayes' Theorem:

@@

\begin{equation}
	p(x_i)=p(x_i|L_i)p(L_i).
\end{equation}
Now, the final step is to solve for the probability distribution of distances, $p(L_i)$. Since we want the probability of absolute differences, we can neglect the distinction between `initial' and `final' and calculate an equilibrium distribution of distances, $p(L)$. To do so, we used a reoccurence relation, where we defined an initial $p^0(L)$ that is not in equilibrium and multiplied it by the conditional probability distribution for arbitrary $L$'s, $p(L'|L)$, until it reaches equilibrium. That is,
\begin{align*}
	p^1(L)= & p(L'|L)p^0(L) \\
	p^2(L)= & p(L'|L)p^1(L)=p(L'|L)^2p^0(L)\\
	&\vdots \\
	p^n(L)= & p(L'|L)p^{n-1}(L)=p(L'|L)^np^0(L) 
\end{align*}
\begin{equation}
	p^n(L)=p(L'|L)^np^0(L) 	
\end{equation}
where $n$ is an arbitrarily large number of steps to ensure the dynein is in equilibrium. This process can be thought of as a large row of dyneins all starting with the probability of an initial distance $p^0(L)$. Once they all take a step, their probability of having another distance $L'$ is $p(L'|L)p(L)$, and after a significant number of $n$ steps, the probability of having a distance $L$ would be the same for all the dyneins, thus being a equilibirum distribution of $L$. To solve for $p(L'|L)$, we can use the previous probability distributions $p(x_f|x_i)$ and $p(x_i|L_i)$ to convert these probabilities into distances, as shown below:
\begin{equation}
	p(L'|L)=I(L'|x_i)p(x_f|x_i)p(x_i|L),
\end{equation}
where $I(L'|x_i)$ is the identity matrix for converting $x_f$ into $L'$. We have now solved for all the unknowns to generate a joint probability distribution of initial and final displacements, $p(x_i,x_f)$.

### 2021-01-31 papers/jin-thesis/main.tex
Added a lot to my thesis directory
Paper edits
Edited thesis by adding refs and adjusted transition matrix part

@@

	\chapter{Introduction}
	\input{./Chapters/1Introduction}
	\chapter{Dynein}
	\input{./Chapters/2Dynein}
	\chapter{Theory}
	\input{./Chapters/3Theory}
	\chapter{Methods}
	\input{./Chapters/4Methods}
	\chapter{Results \& Discussion}
	\input{./Chapters/5Results}
	\chapter{Conclusion}
	\input{./Chapters/6Conclusion}

\newpage
\cite{Burgess2003} \cite{Cianfrocco2015mechanism} \cite{Dewitt2012} \cite{Capek2017}
\bibliographystyle{unsrt}
\bibliography{references}

>>>>

\floatstyle{plaintop}
\newfloat{algorithm}{thb}{lop}
\floatname{algorithm}{Algorithm}
\newenvironment{alg}
{\hrulefill\begin{enumerate}}
{\end{enumerate}\hrulefill}

@@

	\chapter{Introduction}
	\input{./Chapters/1Introduction}
	\chapter{Theory}
	\input{./Chapters/2Theory}
	\chapter{Methods}
	\input{./Chapters/3Methods}
	\chapter{Results \& Discussion}
	\input{./Chapters/4Results}
	\chapter{Conclusion}
	\input{./Chapters/5Conclusion}

\newpage
\cite{Burgess2003} \cite{Cianfrocco2015mechanism} \cite{Dewitt2012} \cite{Capek2017, waczak2019drunken}
\cite{GoodsellArt} \cite{eschbach2011cytoplasmic}
\cite{rao2019molecular} \cite{manna2020mechanistic} \cite{desantis2017lis1} 
\cite{elshenawy2020lis1} \cite{ando2020small} \cite{kinoshita2018step} \cite{qiu2012dynein} \cite{muller1973dynamic} \cite{lim2007vorticity} \cite{fehr2008kinesin} \cite{trott2018mathematical}

\bibliographystyle{unsrt}
\bibliography{references}
